---
title: "Sensor_Plotting"
output: html_document
date: "2025-11-24"
---


## read in packages and data

```{r}
#packages 
library(tidyverse)

### Date levels for plotting
levels_db <- c("22 May", "19 Jun", "11 Jul", "14 Aug", "16 Sep", 
               "30 Sep", "28 Oct", "17 Dec", "26 Feb", "16 Apr")


#read in data
sensors <- read_csv("../Data/sensors_joined.csv") 
 

###filter depths from 0.1 meter resolution for easier plotting viz to 1 meter intervals
high_res_sites <- c(50, 88, 90)   # example — replace with your actual 3 sites

df_filtered <- sensors %>%
  group_by(Site, Date) %>%
  mutate(max_depth = max(Depth_m, na.rm = TRUE)) %>%
  filter(
    # keep all depths for non–high-res sites
    !Site %in% high_res_sites |
    
    # for high-res sites, keep:
    Depth_m == 0.1 |                # the 0.1 m depth
   Depth_m %% 1 == 0 |             # whole meters (1,2,3,...)
   # Depth_m %% 0.5 == 0 |          # half meter intervals 
    Depth_m == max_depth            # the actual deepest measurement
  ) %>%
  ungroup()


```





## Calculate stream water fraction

```{r}
#### calculate stream water fraction adapting methods from Eveleth et al. 2024
##get mean reservoir conductivity by date and make a look up table
Lake_lookup <- sensors |> filter(Site %in% c(88,50)) |> select(Date, SpCond_uScm) |> 
  group_by(Date) |> 
  summarise(Lake_SC = mean(SpCond_uScm, na.rm = T))

##make lookup table for stream cond by sampling date
S1_lookup <- sensors |> filter(Site == 101) |> select(Date, SpCond_uScm) |> 
  rename(S1_SC = SpCond_uScm)


##join look up tables to sensors df for plotting
SWF_data <- df_filtered |> 
  #join data sets
  left_join(S1_lookup, by = "Date") |> 
  left_join(Lake_lookup, by = "Date") |> 
  #Calc stream water fraction
  mutate(SWF = (SpCond_uScm - Lake_SC) / (S1_SC - Lake_SC) ) |> 
  #Calc water density as well (just based on individual measures of water temp) for another metric
  mutate(density = rLakeAnalyzer::water.density(Temp_C)) 



## Plot data for Temp
temp_SWF_s1_c1 <- SWF_data |> 
  #get sites of interest
  filter(!Site %in% c(50,88)) |> 
  #set dry start and end values to NA besides for S1 so shading doesn't overlap
  mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> 
  #format dates
  mutate(Date1 = format(Date, "%d %b")) |> 
  #plot
  ggplot(aes(x = Distance, y = Depth_m, fill = Temp_C, size = SWF))+
  geom_hline(yintercept = 0, color = "blue")+ # line for surface of the water
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #Line for dry reach; color gets rid of border
    geom_line(aes(y = Max_Depth_Site_m+0.1), size = 1, color = "brown")+ #line for max depth
  facet_wrap(~factor(Date1, levels = levels_db), nrow = 1)+
  geom_point(shape = 21)+ #need shape here if not doing site type
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 14),
                    legend.text = element_text(size = 12),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)",
       y = "Depth (m)",
       fill = "Temp (C)")+
  scale_y_reverse(breaks = c(0, 0.5, 1, 1.5, 2, 2.5))+
  scale_x_continuous(breaks = c(0, 150, 300)) +
  scale_fill_distiller(palette = "Spectral", direction = -1) #direction flips value order so red is higher
  
temp_SWF_s1_c1


## Plot data for SpCOND
SC_SWF_s1_c1 <- SWF_data |> 
  #get sites of interest
  filter(!Site %in% c(50,88,90)) |> 
  #set dry start and end values to NA besides for S1 so shading doesn't overlap
  mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> 
  #format dates
  mutate(Date1 = format(Date, "%d %b")) |> 
  #plot
  ggplot(aes(x = Distance, y = Depth_m, fill = SpCond_uScm, size = SWF))+
  geom_hline(yintercept = 0, color = "blue")+ # line for surface of the water
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #Line for dry reach; color gets rid of border
    geom_line(aes(y = Max_Depth_Site_m+0.1), size = 1, color = "brown")+ #line for max depth
  facet_wrap(~factor(Date1, levels = levels_db), nrow = 1)+
  geom_point(shape = 21)+ #need shape here if not doing site type
  theme_bw()+ theme(legend.position = "right", text = element_text(size = 14),
                    legend.text = element_text(size = 12),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)",
       y = "Depth (m)",
       fill = "SC (uS/cm)")+
  scale_y_reverse(breaks = c(0, 0.5, 1, 1.5, 2, 2.5))+
  scale_x_continuous(breaks = c(0, 150, 300)) +
  scale_fill_distiller(palette = "Spectral", direction = -1) #direction flips value order so red is higher

SC_SWF_s1_c1

## Plot data for DO
DO_SWF_s1_c1 <- SWF_data |> 
  #get sites of interest
  filter(!Site %in% c(50,88,90)) |> 
  #set dry start and end values to NA besides for S1 so shading doesn't overlap
  mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> 
  #format dates
  mutate(Date1 = format(Date, "%d %b")) |> 
  #plot
  ggplot(aes(x = Distance, y = Depth_m, fill = DOsat_percent, size = SWF))+
  geom_hline(yintercept = 0, color = "blue")+ # line for surface of the water
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #Line for dry reach; color gets rid of border
    geom_line(aes(y = Max_Depth_Site_m+0.1), size = 1, color = "brown")+ #line for max depth
  facet_wrap(~factor(Date1, levels = levels_db), nrow = 1)+
  geom_point(shape = 21)+ #need shape here if not doing site type
  theme_bw()+ theme(legend.position = "right", text = element_text(size = 14),
                    legend.text = element_text(size = 12),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)",
       y = "Depth (m)",
       fill = "DO (%sat)")+
  scale_y_reverse(breaks = c(0, 0.5, 1, 1.5, 2, 2.5))+
  scale_x_continuous(breaks = c(0, 150, 300)) +
  scale_fill_distiller(palette = "Spectral", direction = -1) #direction flips value order so red is higher

DO_SWF_s1_c1



```

#Function to plot SWF 

```{r}

plot_swf <- function(df,
                     fill_var = NULL,       # variable to fill by
                     fill_var_label = NULL, #what text to use for color ramp label
                     Sites = NULL,          # sites to remove
                     X_breaks = NULL,       # numeric vector
                     Y_breaks = NULL) {     # numeric vector
  
  df %>%
    filter(Site_code %in% Sites) %>%
    mutate(
      Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
      Dry_end   = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA),
      Date1     = format(Date, "%d %b")
    ) %>%
    ggplot(aes(x = Distance, y = Depth_m, fill = {{ fill_var }}, size = SWF)) +
      geom_hline(yintercept = 0, color = "blue") +
      geom_rect(aes(xmin = Dry_start, xmax = Dry_end,
                    ymin = -Inf, ymax = Inf),
                alpha = 1, fill = "gray", color = NA) +
      geom_line(aes(y = Max_Depth_Site_m + 0.1),
                size = 1, color = "brown") +
      facet_wrap(~factor(Date1, levels = levels_db), nrow = 1) +
      geom_point(shape = 21) +
      theme_bw() +
      theme(
        legend.position = "top",
        text = element_text(size = 14),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key = element_blank(),          # remove gray key
        legend.background = element_blank()    # remove gray box behind legend
           ) +
      labs(
        x = "Distance from Stream (m)",
        y = "Depth (m)",
        fill = fill_var_label
      ) +
      scale_y_reverse(breaks = Y_breaks) +
      scale_x_continuous(breaks = X_breaks) +
      scale_size_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1),
                      # range = c(1, 6),   # adjust if you want bigger/smaller points
                      guide = "legend")+
      scale_fill_distiller(palette = "Spectral", direction = -1)
  
}#end function


```


## Make plots of SWF for sites S1 to C1

```{r}
x_breaks_list_S1_C1 <- c(0, 150, 300)
y_breaks_list_S1_C1 <- c(0, 0.5, 1, 1.5, 2, 2.5)

## TEMP
temp_SWF_s1_c1 <- plot_swf(
  df = SWF_data,
  fill_var = Temp_C,
  fill_var_label = "Temp (C)",
  Sites = c("S1", "S2", "B1", "B2", "C1"),
  X_breaks = x_breaks_list_S1_C1,
  Y_breaks = y_breaks_list_S1_C1
)

temp_SWF_s1_c1

## COND
SC_SWF_s1_c1 <- plot_swf(
  df = SWF_data,
  fill_var = SpCond_uScm,
  fill_var_label = "SC (uS/cm)",
  Sites = c("S1", "S2", "B1", "B2", "C1"),
  X_breaks = x_breaks_list_S1_C1,
  Y_breaks = y_breaks_list_S1_C1
)

SC_SWF_s1_c1

```

 
## MS figures 


Stream to C1 for Main MS

```{r}

#### COnd
summary(sensors$SpCond_uScm)

cond_S1_C1 <- sensors |> 
  filter(!Site %in% c(50,88,90)) |> 
    mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> #so just one row has info for dry start and end, was having issues where laying multiple rectangles for more observations affected shading color
  mutate(Date1 = format(Date, "%d %b")) |> 
  ggplot(aes(x = Distance, y = Depth_m, fill = SpCond_uScm))+
  geom_hline(yintercept = 0, color = "blue")+ # line for surface of the water
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #Line for dry reach; color gets rid of border
    geom_line(aes(y = Max_Depth_Site_m+0.1), color = "brown")+ #line for max depth
  facet_wrap(~factor(Date1, levels = levels_db), scales = "fixed", nrow = 1)+
  geom_point(size = 3, shape = 21)+ #need shape here if not doing site type
  scale_y_reverse(breaks = c(0, 0.5, 1, 1.5, 2, 2.5))+
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 12),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)",
       y = "Depth (m)", fill = "SC (uS/cm)")+ # can use \n to break the text
  scale_x_continuous(breaks = c(0, 150, 300)) +
  scale_fill_distiller(palette = "Spectral", direction = -1)

cond_S1_C1

#### Temp
summary(sensors$Temp_C)

temp_S1_C1 <- sensors |> 
  filter(!Site %in% c(50,88,90)) |> 
    mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> #so just one row has info for dry start and end, was having issues where laying multiple rectangles for more observations affected shading color
  mutate(Date1 = format(Date, "%d %b")) |> 
  ggplot(aes(x = Distance, y = Depth_m, fill = Temp_C))+
  geom_hline(yintercept = 0, color = "blue")+ # line for surface of the water
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #Line for dry reach; color gets rid of border
    geom_line(aes(y = Max_Depth_Site_m+0.1), color = "brown")+ #line for max depth
  facet_wrap(~factor(Date1, levels = levels_db), scales = "fixed", nrow = 1)+
  geom_point(size = 3, shape = 21)+ #need shape here if not doing site type
  scale_y_reverse(breaks = c(0, 0.5, 1, 1.5, 2, 2.5))+
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 12),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)",
       y = "Depth (m)", fill = "Temp (C)")+
  scale_x_continuous(breaks = c(0, 150, 300)) +
  scale_fill_distiller(palette = "Spectral", direction = -1)


temp_S1_C1

#### DO
summary(sensors$DOsat_percent)

do_S1_C1 <- sensors |> 
  filter(!Site %in% c(50,88,90)) |> 
    mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> #so just one row has info for dry start and end, was having issues where laying multiple rectangles for more observations affected shading color
  mutate(Date1 = format(Date, "%d %b")) |> 
  ggplot(aes(x = Distance, y = Depth_m, fill = DOsat_percent))+
  geom_hline(yintercept = 0, color = "blue")+ # line for surface of the water
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #Line for dry reach; color gets rid of border
    geom_line(aes(y = Max_Depth_Site_m+0.1), color = "brown")+ #line for max depth
  facet_wrap(~factor(Date1, levels = levels_db), scales = "fixed", nrow = 1)+
  geom_point(size = 3, shape = 21)+ #need shape here if not doing site type
  scale_y_reverse(breaks = c(0, 0.5, 1, 1.5, 2, 2.5))+
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 12),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)",
       y = "Depth (m)", fill = "DO (% sat)")+
  scale_x_continuous(breaks = c(0, 150, 300)) +
scale_fill_distiller(palette = "Spectral", direction = -1)

do_S1_C1


#### Stack toghether
# sensors_S1_C1 <- cond_S1_C1 / temp_S1_C1 / do_S1_C1

sensors_S1_C1 <- cowplot::plot_grid(temp_S1_C1, cond_S1_C1, do_S1_C1, labels = c("a", "b", "c"), ncol = 1)

sensors_S1_C1

# ggsave("../Figures/sensors_S1_C1_fig.png", sensors_S1_C1, width = 11, height = 8, units = "in")


```
 

S1 to P2 By date profiles for Temp, DO, Cond

```{r}
### Date levels
levels_db <- c("22 May", "19 Jun", "11 Jul", "14 Aug", "16 Sep", 
               "30 Sep", "28 Oct", "17 Dec", "26 Feb", "16 Apr")

##COnd 
SpCond_fig <- df_filtered |> 
    mutate(Date1 = format(Date, "%d %b")) |> 
    mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> #so just one row has info for dry start and end, was having issues where laying multiple rectangles for more observations affected shading color
  ggplot(aes(x = Distance, y = Depth_m, color = SpCond_uScm))+
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #color gets rid of border
  facet_wrap(~factor(Date1, levels = levels_db), scales = "fixed", nrow = 1)+
  geom_point(size = 3)+
  scale_y_reverse()+
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 18),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)", y = "Depth (m)", color = "SC (uS/cm)")+
  scale_x_continuous(breaks = c(0, 600, 1200)) +
  scale_color_distiller(palette = "Spectral", direction = -1, breaks = c(25,75,125))

SpCond_fig


##Temp 
Temp_fig <- sensors |> 
    mutate(Date1 = format(Date, "%d %b")) |> 
    filter(!is.na(Temp_C)) |> 
    mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> #so just one row has info for dry start and end, was having issues where laying multiple rectangles for more observations affected shading color
  ggplot(aes(x = Distance, y = Depth_m, color = Temp_C))+
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #color gets rid of border
  facet_wrap(~factor(Date1, levels = levels_db), scales = "fixed", nrow = 1)+
  geom_point(size = 3)+ 
  scale_y_reverse()+
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 18),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)", y = "Depth (m)", color = "Temp (C)")+
  scale_x_continuous(breaks = c(0, 600, 1200)) +
  scale_color_distiller(palette = "Spectral", direction = -1)

Temp_fig


##DO 
DO_fig <- sensors |> 
  filter(!is.na(DOsat_percent)) |> 
    mutate(Date1 = format(Date, "%d %b")) |> 
    mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> #so just one row has info for dry start and end, was having issues where laying multiple rectangles for more observations affected shading color
  ggplot(aes(x = Distance, y = Depth_m, color = DOsat_percent))+
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #color gets rid of border
  facet_wrap(~factor(Date1, levels = levels_db), scales = "fixed", nrow = 1)+
  geom_point(size = 3)+ 
  scale_y_reverse()+
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 18),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)", y = "Depth (m)", color = "DO (% sat)")+
  scale_x_continuous(breaks = c(0, 600, 1200)) +
  scale_color_distiller(palette = "Spectral", direction = -1, breaks = c(25,75,125))

DO_fig

## put them together
library(patchwork)
# sensors_S1_P2 <- Temp_fig / DO_fig / SpCond_fig

sensors_S1_P2 <- cowplot::plot_grid(Temp_fig, SpCond_fig, DO_fig, labels = c("a", "b", "c"), ncol = 1)

sensors_S1_P2
 ggsave("../Figures/sensors_S1_P2_SI_fig.png", sensors_S1_P2, width = 11, height = 8, units = "in")



```

## Extra exploratory figure code 

all variables by date plot 

```{r}

# sensors |> 
#   # filter(!Site %in% c(50,88,90)) |> #If wanting to zoom in shallow sites
#   select(-DO_mgL) |> 
#   pivot_longer(-c(1:9)) |> 
#   # filter(name == "Temp_C") |> 
#   ggplot(aes(x = Distance, y = value, fill = Depth_m,
#              #shape = factor(Site_Class, levels = c("Stream", "Pool", "Cove", "Pelagic"))
#              ))+
#   geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
#             alpha = 0.1, fill = "gray", color = NA )+ #color gets rid of border
#   facet_grid(name~Date, scales = "free_y")+
#   geom_point(size = 3, shape = 21)+ #need shape here if not doing site type
#   theme_bw()+ theme(legend.position = "top", text = element_text(size = 18),
#                     panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
#   labs(x= "Distance from Stream (m)",
#        shape = "Site Type")+
#   scale_x_continuous(breaks = c(0, 500, 1000, 1500))+ #c(250, 750)
#   #scale_shape_manual(values = c(24, 22, 21, 25))+ #18, 15, 16, 17
#   scale_fill_gradient2(low = "red",  high ="blue",
#                        midpoint = 6,  guide = "colourbar", breaks = c(3,6,9, 15, 20))


```

Highlight temp differences by date in shallower sites

```{r}
### Date levels
levels_db <- c("22 May", "19 Jun", "11 Jul", "14 Aug", "16 Sep", 
               "30 Sep", "28 Oct", "17 Dec", "26 Feb", "16 Apr")


##lines 
b1_c1_tempprofiles <- sensors |> 
  filter(!Site %in% c(50,88,90, 100, 101)) |> 
     mutate(Date1 = format(Date, "%d %b")) |> 
  ggplot(aes(x = Temp_C, y = Depth_m, color = factor(Date1, levels = levels_db) ))+
    facet_wrap(~factor(Site_code), scales = "fixed", ncol = 1)+
  geom_point(size =2)+ #need shape here if not doing site type
  geom_line(orientation = "y")+
  scale_y_reverse(breaks = c(0, 0.5, 1, 1.5, 2, 2.5))+
  scale_x_continuous(breaks = c(5,10,15,20,25,30))+
  theme_bw()+ theme(legend.position = "right", text = element_text(size = 14),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Temp (C)", fill = element_blank(),
       y = "Depth (m)", color = "Sampling Date")+
    scale_color_brewer(palette = "Spectral", direction = -1) #direction flips value order so red is higher

b1_c1_tempprofiles

# plotly::ggplotly(b1_c1_tempprofiles)

  


```

## Heatmaps for sites with CTD casts 


make a function 
adapted from: https://github.com/CareyLabVT/Reservoirs/blob/master/Data/DataAlreadyUploadedToEDI/EDIProductionFiles/Heatmap_EDI_function.R

```{r}

# ##### Function #
# heatmap_EDI <- function(data, reservoir, site_code, z){
#   
#   # load packages
#   pacman::p_load(tidyverse, lubridate, akima, reshape2, 
#                gridExtra, grid, colorRamps, RColorBrewer, cowplot, rLakeAnalyzer)
#   
#   #find thermocline data
# thermocline_data <- data |> 
#   filter(Site_code == site_code) |> 
#   select(Date, Depth_m, Temp_C) |> 
#   pivot_wider(names_from = Depth_m, values_from = Temp_C, values_fn = list(Temp_C = mean)) |> 
#   rename(datetime = Date)
# 
# colnames(thermocline_data)[-1] = paste0('wtr_',colnames(thermocline_data)[-1])
# 
# 
# thermocline_depths <- ts.thermo.depth(thermocline_data, na.rm = TRUE)
# 
# thermocline_depths <- thermocline_depths |> 
#   rename(Date = datetime)
#   
#   #subset to relevant data
#   fp_new <- data %>%
#     filter(Reservoir %in% reservoir & Site_code %in% site_code) %>%
#     select(Date, Depth_m, {{z}}) 
#   
#   
#   # Create title for plot
#   fig_title <- paste(reservoir, "Site", site_code, z, sep = " ")
#   
#   interp <- interp(x=as.numeric(fp_new$Date), y = fp_new$Depth_m, z = unlist(fp_new[z]),
#                    #xo = seq(min(fp_new$DOY), max(fp_new$DOY), by = .1), 
#                    xo = seq(min(fp_new$Date), max(fp_new$Date), by = "day"),
#                    yo = seq(min(fp_new$Depth_m), max(fp_new$Depth_m), by = 0.1),
#                    extrap = T, linear = T, duplicate = "strip")
#   interp <- interp2xyz(interp, data.frame=T)
#   
#   interp<-interp%>%
#     mutate(Date=as.Date(x, origin = "1970-01-01"))
#   
#   
#   p1 <- ggplot(interp, aes(x=Date, y=y))+
#     geom_raster(aes(fill=z))+
#     scale_y_reverse(expand = c(0,0))+
#     scale_x_date(expand = c(0, 0)) +
#     ## thermocline lines
#   geom_line(data = thermocline_depths, aes(x=Date, y=thermo.depth, z=NULL), color = "black", lwd = 1)+
#   geom_point(data = thermocline_depths, aes(x=Date, y=thermo.depth, z=NULL), pch = 21, size = 2, color = "white", fill = "black")+
#     ##sampling points
#     geom_point(data = fp_new, aes(x = Date, y = 0.1, z = NULL), pch = 25, size = 3, fill = "black")+ 
#     scale_fill_gradientn(colours = blue2green2red(60), na.value="gray")+
#     labs(x = "Date", y = "Depth (m)", title = fig_title,fill=(z))+
#     theme_bw()
#   
#   return(p1)
#   
# }

 
 
# #### Site P2 heatmap
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "P2", z = "Temp_C")
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "P1", z = "Temp_C")
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "C2", z = "Temp_C")
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "P2", z = "SpCond_uScm")
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "P1", z = "SpCond_uScm")
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "C2", z = "SpCond_uScm")
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "P2", z = "DO_mgL")
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "P1", z = "DO_mgL")
# heatmap_EDI(data = sensors, reservoir = "CCR", site_code = "C2", z = "DO_mgL")


```

## calc thermocline by dates

```{r}
# library(rLakeAnalyzer)
# 
# ## P2 thermocline
# p2_temp <- sensors |> 
#   filter(Site_code == "P2") |> 
#   select(Date, Depth_m, Temp_C) |> 
#   pivot_wider(names_from = Depth_m, values_from = Temp_C, values_fn = list(Temp_C = mean)) |> 
#   rename(datetime = Date)
# 
# colnames(p2_temp)[-1] = paste0('wtr_',colnames(p2_temp)[-1])
# 
# p2_thermocline <- ts.thermo.depth(p2_temp, na.rm = TRUE)
# 
# ## P1 thermocline
# p1_temp <- sensors |> 
#   filter(Site_code == "P1") |> 
#   select(Date, Depth_m, Temp_C) |> 
#   pivot_wider(names_from = Depth_m, values_from = Temp_C, values_fn = list(Temp_C = mean)) |> 
#   rename(datetime = Date)
# 
# colnames(p1_temp)[-1] = paste0('wtr_',colnames(p1_temp)[-1])
# 
# p1_thermocline <- ts.thermo.depth(p1_temp, na.rm = TRUE)
# 
# 
# ## C2 thermocline
# c2_temp <- sensors |> 
#   filter(Site_code == "C2") |> 
#   select(Date, Depth_m, Temp_C) |> 
#   pivot_wider(names_from = Depth_m, values_from = Temp_C, values_fn = list(Temp_C = mean)) |> 
#   rename(datetime = Date)
# 
# colnames(c2_temp)[-1] = paste0('wtr_',colnames(c2_temp)[-1])
# 
# c2_thermocline <- ts.thermo.depth(c2_temp, na.rm = TRUE)



```







