---
title: "Sensor_Plotting"
output: html_document
date: "2025-11-24"
---


## read in packages and data

```{r}
#packages 
library(tidyverse)
library(rLakeAnalyzer) #water.density

#read in data
sensors <- read_csv("../Data/sensors_joined.csv") 
 

###filter depths from 0.1 meter resolution for easier plotting viz to 1 meter intervals
high_res_sites <- c(50, 88, 90)   # example — replace with your actual 3 sites

sensors_filt <- sensors %>%
  group_by(Site, Date) %>%
  mutate(max_depth = max(Depth_m, na.rm = TRUE)) %>%
  filter(
    # keep all depths for non–high-res sites
    !Site %in% high_res_sites |
    
    # for high-res sites, keep:
    Depth_m == 0.1 |                # the 0.1 m depth
   # Depth_m %% 1 == 0 |             # whole meters (1,2,3,...)
    Depth_m %% 0.5 == 0 |          # half meter intervals 
    Depth_m == max_depth            # the actual deepest measurement
  ) %>%
  ungroup()


```


## Make levels for plotting 

```{r}
### Date levels for plotting
levels_db <- c("22 May", "19 Jun", "11 Jul", "14 Aug", "16 Sep", 
               "30 Sep", "28 Oct", "17 Dec", "26 Feb", "16 Apr")


## set axis break levels for different subsets
#S1 to C1
x_breaks_list_S1_C1 <- c(0, 150, 300)
y_breaks_list_S1_C1 <- c(0, 0.5, 1, 1.5, 2, 2.5)

#s1 to P2
x_breaks_list_S1_P2 <- c(0, 500, 1000, 1500)
y_breaks_list_S1_P2 <- c(0, 3, 6, 9, 12, 15, 18, 21)



```





## Plot just sensor data

```{r}

plot_sensors <- function(df,
                     fill_var = NULL,       # variable to fill by
                     fill_var_label = NULL, # what text to use for color ramp label
                     Sites = NULL,          # sites to remove
                     X_breaks = NULL,       # numeric vector
                     Y_breaks = NULL       # numeric vector
                     ) {     
  
  df %>%
    filter(Site_code %in% Sites) %>%
    mutate(
      Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
      Dry_end   = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA),
      Date1     = format(Date, "%d %b")
    ) %>%
    ggplot(aes(x = Distance, y = Depth_m, fill = {{ fill_var }} )) +
      geom_hline(yintercept = 0, color = "blue") +
      geom_rect(aes(xmin = Dry_start, xmax = Dry_end,
                    ymin = -Inf, ymax = Inf),
                alpha = 1, fill = "gray", color = NA) +
      geom_line(aes(y = Max_Depth_Site_m + 0.1),
                size = 1, color = "brown") +
      facet_wrap(~factor(Date1, levels = levels_db), nrow = 1) +
      geom_point(size = 3, shape = 21) +
      theme_bw() +
      theme(
        legend.position = "top",
        text = element_text(size = 14),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ) +
      labs(
        x = "Distance from Stream (m)",
        y = "Depth (m)",
        fill = fill_var_label
      ) +
      scale_y_reverse(breaks = Y_breaks) +
      scale_x_continuous(breaks = X_breaks) +
      scale_fill_distiller(palette = "Spectral", direction = -1)
  
} # end function


```


 
Use function to plot sensors S1 to C1

```{r}

## TEMP
plot_sensors(
  df = sensors_filt,
  fill_var = Temp_C,
  fill_var_label = "Temp (C)",
  Sites = c("S1", "S2", "B1", "B2", "C1"),
  X_breaks = x_breaks_list_S1_C1,
  Y_breaks = y_breaks_list_S1_C1
)

```

 
 Use function to plot sensors S1 to P2

```{r}

## TEMP
plot_sensors(
  df = sensors_filt,
  fill_var = Temp_C,
  fill_var_label = "Temp (C)",
  Sites = c("S1", "S2", "B1", "B2", "C1", "C2", "P1", "P2"),
  X_breaks = x_breaks_list_S1_P2,
  Y_breaks = y_breaks_list_S1_P2
)

```
 
 
 
## OLD MS figures 


Stream to C1 for Main MS

```{r}

#### COnd
summary(sensors$SpCond_uScm)

sensors |> 
  filter(!Site %in% c(50,88,90)) |> 
    mutate(Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
         Dry_end = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA)) |> #so just one row has info for dry start and end, was having issues where laying multiple rectangles for more observations affected shading color
  mutate(Date1 = format(Date, "%d %b")) |> 
  ggplot(aes(x = Distance, y = Depth_m, fill = SpCond_uScm))+
  geom_hline(yintercept = 0, color = "blue")+ # line for surface of the water
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 1, fill = "gray", color = NA )+ #Line for dry reach; color gets rid of border
    geom_line(aes(y = Max_Depth_Site_m+0.1), color = "brown")+ #line for max depth
  facet_wrap(~factor(Date1, levels = levels_db), scales = "fixed", nrow = 1)+
  geom_point(size = 3, shape = 21)+ #need shape here if not doing site type
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 12),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)",
       y = "Depth (m)", fill = "SC (uS/cm)")+ # can use \n to break the text
  scale_x_continuous(breaks = c(0, 150, 300), limits = c(-40, 380)) +
  ylim(-0.1,3)+
  scale_y_reverse(breaks = c(0, 0.5, 1, 1.5, 2, 2.5), 
                  #limits = c(0.2, 3)
                  )+
  ##add labels for site type starting
     annotate("text",
           x = c(0, 189, 336),
           y = -0.1,
           label = c("S1", "B1", "C1"),
           # hjust = -0.01,     # pushes labels above the plotting area
           size = 3
           )+
  scale_fill_distiller(palette = "Spectral", direction = -1)




```
 
 
 
 
 
 ----------------------------------- SWF --------------------------------------------------------
 
 
 
 
## Calculate stream water fraction

```{r}
#### calculate stream water fraction adapting methods from Eveleth et al. 2024
##get mean reservoir conductivity by date and make a look up table
Lake_lookup <- sensors_filt |> filter(Site %in% c(50)) |> dplyr::select(Date, SpCond_uScm) |> 
  group_by(Date) |> 
  summarise(Lake_SC = mean(SpCond_uScm, na.rm = T))

##make lookup table for stream cond by sampling date
S1_lookup <- sensors_filt |> filter(Site == 101) |> dplyr::select(Date, SpCond_uScm) |> 
  rename(S1_SC = SpCond_uScm)


##join look up tables to sensors df for plotting
SWF_data <- sensors_filt |> 
  #join data sets
  left_join(S1_lookup, by = "Date") |> 
  left_join(Lake_lookup, by = "Date") |> 
  #Calc water density as well (just based on individual measures of water temp) for another metric
  mutate(density = water.density(Temp_C)) |>
  #Calc stream water fraction
  mutate(SWF = (SpCond_uScm - Lake_SC) / (S1_SC - Lake_SC) ) 
  


```
 
 
## Function to plot SWF 

```{r}

plot_swf <- function(df,
                     fill_var = NULL,       # variable to fill by
                     fill_var_label = NULL, # what text to use for color ramp label
                     Sites = NULL,          # sites to remove
                     X_breaks = NULL,       # numeric vector
                     Y_breaks = NULL,       # numeric vector
                     show_depthlines = TRUE   # toggle blue hline + brown geom_line
                      ) {
  
  p <- df %>%
    filter(Site_code %in% Sites) %>%
    mutate(
      Dry_start = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_start, NA),
      Dry_end   = ifelse(Depth_m == 0.1 & Site_code == "S1", Dry_end, NA),
      Date1     = format(Date, "%d %b")
    ) %>%
    ggplot(aes(x = Distance, y = Depth_m, fill = {{ fill_var }}, size = SWF))
  
  if (show_depthlines) {
    p <- p + geom_hline(yintercept = 0, color = "blue")
  }
  
  p <- p +
    geom_rect(aes(xmin = Dry_start, xmax = Dry_end,
                  ymin = -Inf, ymax = Inf),
              alpha = 1, fill = "gray", color = NA)
  
  if (show_depthlines) {
    p <- p + geom_line(aes(y = Max_Depth_Site_m + 0.1), size = 1, color = "#654321")
  }
  
  p <- p +
    facet_wrap(~factor(Date1, levels = levels_db), nrow = 1) +
    geom_point(shape = 21) +
    theme_bw() +
    theme(
      legend.position = "top",
      text = element_text(size = 14),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    ) +
    labs(
      x = "Distance from Stream (m)",
      y = "Depth (m)",
      fill = fill_var_label
    ) +
    scale_y_reverse(breaks = Y_breaks) +
    scale_x_continuous(breaks = X_breaks) +
    scale_fill_distiller(palette = "Spectral", direction = -1)
  
  return(p)
  
} # end function


```
 
 
 
## Make plots of SWF for sites S1 to C1

```{r}

## TEMP
temp_SWF_s1_c1 <- plot_swf(df = SWF_data,
                           fill_var = Temp_C,  fill_var_label = "Temp (C)",
                           Sites = c("S1", "S2", "B1", "B2", "C1"),
                           X_breaks = x_breaks_list_S1_C1,  Y_breaks = y_breaks_list_S1_C1,
                           show_depthlines = TRUE)

temp_SWF_s1_c1

## SC
SC_SWF_s1_c1 <- plot_swf(df = SWF_data,
                         fill_var = SpCond_uScm,  fill_var_label = "SC (uS/cm)",
                         Sites = c("S1", "S2", "B1", "B2", "C1"),
                         X_breaks = x_breaks_list_S1_C1,  Y_breaks = y_breaks_list_S1_C1,
                         show_depthlines = TRUE)

SC_SWF_s1_c1

## DO
DO_SWF_s1_c1 <- plot_swf( df = SWF_data,
                          fill_var = DOsat_percent,  fill_var_label = "DO (%sat)",
                          Sites = c("S1", "S2", "B1", "B2", "C1"),
                          X_breaks = x_breaks_list_S1_C1,  Y_breaks = y_breaks_list_S1_C1,
                           show_depthlines = TRUE)

DO_SWF_s1_c1

##stack plots

sensors_SWF_S1_C1 <- cowplot::plot_grid(temp_SWF_s1_c1, SC_SWF_s1_c1, DO_SWF_s1_c1, labels = c("a", "b", "c"), ncol = 1)

sensors_SWF_S1_C1

# ggsave("../Figures/sensors_SWF_S1_C1_fig.png", sensors_SWF_S1_C1, width = 11, height = 8, units = "in")


```



## Make plots of SWF for sites S1 to P2

```{r}

## TEMP
temp_SWF_s1_p2 <- plot_swf(df = SWF_data,
                           fill_var = Temp_C,  fill_var_label = "Temp (C)",
                           Sites = c("S1", "S2", "B1", "B2", "C1", "C2", "P1", "P2"),
                          X_breaks = x_breaks_list_S1_P2,  Y_breaks = y_breaks_list_S1_P2,
                           show_depthlines = F)

temp_SWF_s1_p2

## SC
SC_SWF_s1_p2 <- plot_swf(df = SWF_data,
                         fill_var = SpCond_uScm,  fill_var_label = "SC (uS/cm)",
                         Sites = c("S1", "S2", "B1", "B2", "C1", "C2", "P1", "P2"),
                          X_breaks = x_breaks_list_S1_P2,  Y_breaks = y_breaks_list_S1_P2,
                           show_depthlines = F)

SC_SWF_s1_p2

## DO
DO_SWF_s1_p2 <- plot_swf( df = SWF_data,
                          fill_var = DOsat_percent,  fill_var_label = "DO (%sat)",
                          Sites = c("S1", "S2", "B1", "B2", "C1", "C2", "P1", "P2"),
                          X_breaks = x_breaks_list_S1_P2,  Y_breaks = y_breaks_list_S1_P2,
                           show_depthlines = F)

DO_SWF_s1_p2

##stack plots

sensors_SWF_S1_C1 <- cowplot::plot_grid(temp_SWF_s1_c1, SC_SWF_s1_c1, DO_SWF_s1_c1, labels = c("a", "b", "c"), ncol = 1)

sensors_SWF_S1_C1

# ggsave("../Figures/sensors_SWF_S1_C1_fig.png", sensors_SWF_S1_C1, width = 11, height = 8, units = "in")


```
 
 
 
 
 
 
 
 
 
 
## SWF using d-excess
same approach as above but now d-excess instead of SC
```{r}

# ## read in chem data
# chem <- read_csv("../Data/chemistry_joined_sensor_hydro.csv")
# 
# #### calculate stream water fraction adapting methods from Eveleth et al. 2024
# ##get mean reservoir d-excess by date and make a look up table
# Lake_d_lookup <- chem |> filter(Site %in% c(50)) |> dplyr::select(Date, D_excess) |> 
#   group_by(Date) |> 
#   summarise(Lake_d_excess = mean(D_excess, na.rm = T))
# 
# ##make lookup table for stream cond by sampling date
# S1_d_lookup <- chem |> filter(Site == 101) |> dplyr::select(Date, D_excess) |> 
#   rename(S1_d_excess = D_excess)
# 
# 
# ##join look up tables to sensors df for plotting
# SWF_Dexcess_data <- chem |> 
#   #join data sets
#   left_join(S1_d_lookup, by = "Date") |> 
#   left_join(Lake_d_lookup, by = "Date") |> 
#   #Calc stream water fraction
#   mutate(SWF = (D_excess - Lake_d_excess) / (S1_d_excess - Lake_d_excess) ) 

```


look at data
```{r}

#  plot_swf(
#   df = SWF_Dexcess_data,
#   fill_var = D_excess,
#   fill_var_label = "d-excess",
#   Sites = c("S1", "S2", "B1", "B2", "C1", "C2", "P1", "P2"),
#   X_breaks = x_breaks_list_S1_P2,
#   Y_breaks = y_breaks_list_S1_P2
# )

```

 
 
 
 
 
 
 
 
 
 
 
 
