---
title: "Breakpoints"
output: html_document
date: "2025-07-21"
---

## load packages and read in data

```{r}
## packages 
library(segmented) #for breakpoints
library(tidyverse)
library(patchwork) # so you can arrange figures by "/" or "|"

## read in chem data
chem <- read_csv("../Data/chemistry_joined.csv")   

## read in chl-a data
ccr_chla <- read.csv("../Data/CCR_chla.csv") |> 
  mutate(Date = as.Date(Date))

chem_chla <- full_join(chem, ccr_chla, by = c("Reservoir", "Site", "Date", "Depth_m")) |> 
  select(-Pheo_ugL)


##read in sensors and align to rows 
sensors <- read_csv("../Data/sensors_joined.csv") |> 
  select(-c(Site_code, Dry_start, Dry_end, Distance))

chem_sensors <- chem_chla %>%
  left_join(
    sensors,
    by = c("Reservoir", "Site", "Date"),
    suffix = c("_chem", "_sens")
  ) %>%
  mutate(depth_diff = abs(Depth_m_chem - Depth_m_sens)) %>%
  group_by(Reservoir, Site, Date, Distance, Depth_m_chem) %>%
  slice_min(depth_diff, n = 1, with_ties = FALSE) %>%
  ungroup() |> 
  rename(Depth_m = Depth_m_chem) |> #set name back for function to work
  #get rid of sensor values for site P2 cast on 19 June that didn't reach the bottom (difference in depth was 7.5 meters)
  mutate(SpCond_uScm = ifelse(Date == ymd("2024-06-19") & Site_code == "P2" & Depth_m == 20, NA, SpCond_uScm),
         Temp_C = ifelse(Date == ymd("2024-06-19") & Site_code == "P2" & Depth_m == 20, NA, Temp_C),
         DO_mgL = ifelse(Date == ymd("2024-06-19") & Site_code == "P2" & Depth_m == 20, NA, DO_mgL),
         DOsat_percent = ifelse(Date == ymd("2024-06-19") & Site_code == "P2" & Depth_m == 20, NA, DOsat_percent)
         ) 


```


## singular breakpoint

```{r}
### BIX as an example

#fit normal linear regression 
my.lm <- lm(BIX ~ Distance, data = chem)
summary(my.lm)
davies <- davies.test(my.lm, k=10)
davies

#Run breakpoint
my.seg <- segmented(my.lm,  seg.Z = ~ Distance,  psi = 2 )
summary(my.seg)
my.seg$psi
slope(my.seg)

#fit breakpoint segments to distance data
my.fitted <- fitted(my.seg)
my.model <- data.frame(Distance = chem$Distance, BIX_fit = my.fitted)

#get breakpoint lines
#NOTE that this line comes up chopy because its fitting to x axis values that aren't continous (have gaps between c and P sites; fixed later below)
my.lines <- my.seg$psi[ , 2]

#plot
chem |> 
  ggplot(aes(x = Distance, y = BIX, fill = Depth_m))+
  geom_point(shape = 21, size = 4) + #shape = 21,
  # geom_smooth()+
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 18),
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)", y = "BIX", fill = "Depth (m)",
       shape = "Site Type")+
  scale_x_continuous(breaks = c(0, 500, 1000, 1500))+ 
  scale_fill_gradient2(low = "red",  high ="blue",
                       midpoint = 6,  guide = "colourbar", breaks = c(3,6,9, 15))+
  #geom_smooth()+
  geom_vline(xintercept = my.lines, linetype = "dashed", size = 1.5) +
  geom_line(data = my.model, aes(x = Distance, y = BIX_fit), size = 1, inherit.aes = FALSE ) 



## get the quick stats by running on variables breakpoint
# #BIX
# my.lm <- lm(BIX ~ Distance, data = chem)
# davies <- davies.test(my.lm, k=10)
# davies
# my.seg <- segmented(my.lm,  seg.Z = ~ Distance,  psi = 2 )
# summary(my.seg)
# my.seg$psi
# slope(my.seg)


```


## Breakpoint function for MS figures

```{r}
library(tidyverse)
library(segmented)
library(rlang)

run_breakpoint_analysis <- function(df, response_var, 
                                    fill_var = NULL, #Variable fill, could also set to "Depth_m"
                                    fill_var_label = NULL #label to assign for color ramp 
                                    ) {
  
  # Convert response variable to symbol
  y <- sym(response_var)
  
  # Fit linear model
  my.lm <- lm(as.formula(paste(response_var, "~ Distance")), data = df)
  
  # Davies test
  davies <- davies.test(my.lm, k = 10)
  
  # Segmented model
  my.seg <- segmented(my.lm, seg.Z = ~ Distance, psi = 2)
  
  # Make regression lines; making a continuous distance gradient by 1 meter interval so lines don't jump like above
  dist_gradient <- data.frame(Distance = seq(0, max(df$Distance, na.rm = TRUE), by = 1))
  dist_gradient$fit <- predict(my.seg, newdata = dist_gradient)
  
  # Breakpoint + SE
  my.lines <- my.seg$psi[, 2]
  my.lines.SE <- my.seg$psi[, 3]
  
  # Build plot
  p <- df |>
    ggplot(aes(x = Distance, y = !!y, fill = .data[[fill_var]])) +
    geom_rect(
      aes(
        xmin = (my.lines - my.lines.SE),
        xmax = (my.lines + my.lines.SE),
        ymin = -Inf, ymax = Inf
      ),
      alpha = 0.1, fill = "gray", color = NA
    ) +
    geom_point(shape = 21, size = 4) +
    labs(
      x = "Distance from Stream (m)",
      y = paste0(response_var),
      fill = fill_var_label
    ) +
        scale_x_continuous(breaks = c(0, 500, 1000, 1500)) +
    ##For other variable color ramping 
    scale_fill_distiller(palette = "Spectral", direction = -1)+
    ##For Depth only plots
    # scale_fill_gradient2(
    #   low = "red", high = "blue", midpoint = 6, #could do median(df[[fill_var]], na.rm = TRUE)
    #   guide = "colourbar"
    # ) +
    geom_vline(xintercept = my.lines, linetype = "dashed", size = 1.5) +
    geom_line(
      data = dist_gradient,
      aes(x = Distance, y = fit),
      size = 1,
      inherit.aes = FALSE
    )+
    ##add labels for site type starting
     annotate("text",
           x = c(0, 189, 336, 1013),
           y = Inf,
           label = c("S1", "B1", "C1", "P1"),
           vjust = -0.5,     # pushes labels above the plotting area
           size = 5) +
  coord_cartesian(clip = "off") +   # allow drawing outside panel
  theme_bw() +
  theme(
      legend.position = "top",
      text = element_text(size = 14),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
        legend.background = element_blank(),
    legend.box.background = element_blank()
    )
 
  # Return everything as a list
  list(
    plot = p,
    segmented_summary = summary(my.seg),
    segmented_slopes = slope(my.seg),
    davies_test = davies
  )
} #end of function


```


## Breakpoints for stream-res boundary 
Temp, SpCond, d-excess

```{r}
## Temp
temp_break <- run_breakpoint_analysis(chem_sensors, response_var = "Temp_C",
                                      fill_var = "Depth_m", fill_var_label = "Depth (m)")

temp_break$plot
temp_break$davies_test

## Cond
SC_break <- run_breakpoint_analysis(chem_sensors, response_var = "SpCond_uScm",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

SC_break$plot
SC_break$davies_test

## DO
DO_break <- run_breakpoint_analysis(chem_sensors, response_var = "DOsat_percent",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

DO_break$plot
DO_break$davies_test


## d-excess
Dexcess_break <- run_breakpoint_analysis(chem_sensors, response_var = "D_excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

Dexcess_break$plot
Dexcess_break$davies_test


## stack Sp cond and d-excess
Dexcess_break$plot / SC_break$plot



```


Check if this changes by days with dry streams or days with plunge points

Set up filtering for day groups 
```{r}
### dry stream vs connected days 
dry_stream_days <- c(ymd("2024-06-19"), ymd("2024-07-11"), ymd("2024-08-14"), 
                ymd("2024-09-16"), ymd("2024-10-28") )

wet_stream_days <- c(ymd("2024-05-22"), ymd("2024-09-30"), ymd("2024-12-17"), 
                ymd("2025-02-26"), ymd("2025-04-16"))

### plunge v no plunge groups 
plungedays <- c(ymd("2024-05-22"), ymd("2024-06-19"), ymd("2025-02-26"), ymd("2025-04-16"))

nonplungedays <- c(ymd("2024-07-11"), ymd("2024-08-14"), ymd("2024-09-16"),
                    ymd("2024-09-30"),  ymd("2024-10-28"), ymd("2024-12-17"))


```


Plunge vs No plunge BP's
```{r}
## SC
plungedays_df <- chem_sensors |> 
  filter(Date %in% plungedays)

SC_break_plunge <- run_breakpoint_analysis(plungedays_df, response_var = "SpCond_uScm",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

SC_break_plunge$davies_test


noplungedays_df <- chem_sensors |> 
  filter(Date %in% nonplungedays)

SC_break_plungeno <- run_breakpoint_analysis(noplungedays_df, response_var = "SpCond_uScm",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

SC_break_plungeno$davies_test

SC_break_plunge$plot / SC_break_plungeno$plot


##D-excess
Dexcess_break_plunge <- run_breakpoint_analysis(plungedays_df, response_var = "D_excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

Dexcess_break_plunge$davies_test


Dexcess_break_plungeno <- run_breakpoint_analysis(noplungedays_df, response_var = "D_excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

Dexcess_break_plungeno$davies_test

Dexcess_break_plungeno$plot / Dexcess_break_plungeno$plot


```

Dry vs Wet  BP's
```{r}
## SC
drystream_df <- chem_sensors |> 
  filter(Date %in% dry_stream_days)

SC_break_dry <- run_breakpoint_analysis(drystream_df, response_var = "SpCond_uScm",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

SC_break_dry$davies_test


wetstream_df <- chem_sensors |> 
  filter(Date %in% wet_stream_days)

SC_break_wet <- run_breakpoint_analysis(wetstream_df, response_var = "SpCond_uScm",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

SC_break_wet$davies_test

SC_break_dry$plot / SC_break_wet$plot


##D-excess
D_break_dry <- run_breakpoint_analysis(drystream_df, response_var = "D_excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

D_break_wet <- run_breakpoint_analysis(wetstream_df, response_var = "D_excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

D_break_dry$plot / D_break_wet$plot



```


## Z score values for Breakpoints

```{r}
chem_ZT <- chem_sensors |> 
  mutate(BIX_Zscore = scale(BIX),
         HIX_Zscore = scale(HIX),
         SUVA254_Zscore = scale(SUVA254),
         AT_Zscore = scale(A_T),
         DOC_Zscore = scale(DOC_mgL),
         NO3_Zscore = scale(NO3NO2_ugL),
         Chla_Zscore = scale(Chla_ugL)
         )

```



#### DOM Breakpoints for MS

set up surface and max depth DFs

```{r}
# ##make surface only data frame
# chem_surf <- chem_sensors |> 
#   filter(Depth_m == 0.1)
# 
# ##make max depth only data frame
# max_depth_df <- chem_sensors |> 
#   group_by(Site, Date) |> 
#   mutate(max_depth = if_else(Depth_m == max(Depth_m, na.rm = TRUE), "Max", "Not")) |> 
#   ungroup() |> 
#   filter(max_depth == "Max")

```

BIX z-score test 

```{r}

chem_z <- chem_sensors |> 
  mutate(BIX_zt = scale(BIX))


bix_zt_break <- run_breakpoint_analysis(chem_ZT, response_var = "BIX_Zscore", 
                                        fill_var = "SpCond_uScm", fill_var_label = "d-excess")
bix_zt_break$plot
bix_zt_break$davies_test
bix_zt_break$segmented_summary
bix_zt_break$segmented_slopes



```

## BIX to lots of responses for fill

```{r}

```







## Breakpoints from MS from early feb 2026
BIX breakpoints: all depths, surface, bottom
To interpret stats outputs its easier to run chunk line by line
```{r}
#### BIX 

## BIX across all of chem
BIX_break <- run_breakpoint_analysis(chem, "BIX")
#see plot
BIX_break$plot
#get stats for MS table
BIX_break$davies_test
BIX_break$segmented_summary
BIX_break$segmented_slopes


## BIX surface
BIX_surf_break <- run_breakpoint_analysis(chem_surf, "BIX")
BIX_surf_break$davies_test
BIX_surf_break$segmented_summary
BIX_surf_break$segmented_slopes

## BIX BOT
BIX_bot_break <- run_breakpoint_analysis(max_depth_df, "BIX")
BIX_bot_break$davies_test
BIX_bot_break$segmented_summary
BIX_bot_break$segmented_slopes


#plot breakpoints
BIX_break$plot | BIX_surf_break$plot | BIX_bot_break$plot


```


DOC breakpoints: all depths, surface, bottom 

```{r}
#### DOC 

## DOC across all of chem
DOC_break <- run_breakpoint_analysis(chem, "DOC_mgL")
DOC_break$davies_test
DOC_break$segmented_summary
DOC_break$segmented_slopes

## DOC surface
DOC_surf_break <- run_breakpoint_analysis(chem_surf, "DOC_mgL")
DOC_surf_break$davies_test
DOC_surf_break$segmented_summary
DOC_surf_break$segmented_slopes

## DOC BOT
DOC_bot_break <- run_breakpoint_analysis(max_depth_df, "DOC_mgL")
DOC_bot_break$davies_test
DOC_bot_break$segmented_summary
DOC_bot_break$segmented_slopes


```

DOC / BIX figure for MS 

```{r}
##plot them all
Breakpoints_MS <- cowplot::plot_grid(DOC_break$plot, BIX_break$plot,  
                                     DOC_surf_break$plot,  BIX_surf_break$plot, 
                                     DOC_bot_break$plot, BIX_bot_break$plot,
                                     labels = c("a", "b", "c", "d", "e", "f"), ncol = 2)

Breakpoints_MS

#ggsave("../Figures/DOM_breakpoints_MS.png", Breakpoints_MS, height = 8, width = 8, units = "in")


```



Breakponts for SI figure
SI breakpoints across all depths
```{r}
#### AT
AT_break <- run_breakpoint_analysis(chem, "A_T")
AT_break$davies_test
AT_break$segmented_summary
AT_break$segmented_slopes


#### HIX
HIX_break <- run_breakpoint_analysis(chem, "HIX")
HIX_break$davies_test
HIX_break$segmented_summary
HIX_break$segmented_slopes

#### SUVA
SUVA_break <- run_breakpoint_analysis(chem, "SUVA254")
SUVA_break$davies_test
SUVA_break$segmented_summary
SUVA_break$segmented_slopes


####NO3
NO3_break <- run_breakpoint_analysis(chem, "NO3NO2_ugL")
NO3_break$davies_test
NO3_break$segmented_summary
NO3_break$segmented_slopes


#### make figure
Breakpoints_SI <- cowplot::plot_grid(AT_break$plot, SUVA_break$plot, HIX_break$plot,  NO3_break$plot, 
                                     labels = c("a", "b", "c", "d"), ncol = 1)

Breakpoints_SI

# ggsave("../Figures/DOM_NO3_breakpoints_SI.png", Breakpoints_SI, height = 9, width = 7, units = "in")

```


## Check out adding chla breakpoint 

run breakpoint 

```{r}
chla_break <- chem_sensors |> 
  filter(!is.na(Chla_ugL))


chla_break_surf <- run_breakpoint_analysis(chla_break, response_var = "Chla_ugL")

chla_break_surf$plot
chla_break_surf$davies_test
chla_break_surf$segmented_summary
chla_break_surf$segmented_slopes


##NO3 surface
no3_break_surf <- run_breakpoint_analysis(chem_surf, response_var = "NO3NO2_ugL")

no3_break_surf$plot
no3_break_surf$davies_test
no3_break_surf$segmented_summary
no3_break_surf$segmented_slopes

##make figure 
chla_break_plots <- cowplot::plot_grid(no3_break_surf$plot, chla_break_surf$plot, 
                                       DOC_surf_break$plot,  BIX_surf_break$plot, 
                                     labels = c("a", "b", "c", "d"), ncol = 1)

chla_break_plots

#ggsave("../Figures/Chla_breakpoints_surf_SI.png", chla_break_plots, height = 9, width = 7, units = "in")


```




## regressions to sensors 

```{r}
zz |> 
  mutate(Site_Type = ifelse(Site_code.x %in% c("S1", "S2"), "Stream", NA),
         Site_Type = ifelse(Site_code.x %in% c("B1", "B2"), "Backwater", Site_Type),
         Site_Type = ifelse(Site_code.x %in% c("C1", "C2"), "Cove", Site_Type),
         Site_Type = ifelse(Site_code.x %in% c("P1", "P2"), "Pelagic", Site_Type)
         ) |> 
  ggplot(aes(x = Temp_C, y = BIX, shape = Site_Type, color = Date))+
  geom_point()



```




