---
title: "Breakpoints"
output: html_document
date: "2025-07-21"
---

## load packages and read in data

```{r}
## packages 
library(segmented) #for breakpoints
library(tidyverse)
library(patchwork) # so you can arrange figures by "/" or "|"

## read in chem data
chem <- read_csv("../Data/chemistry_joined_sensor_hydro.csv")   

```


## Set up data for breakpoints

z-scores and make depth bins

```{r}
chem_bp <- chem |> 
  #Make depth column grouped by Surface, Bottom, and Intermediate (just for P1 and P2 for INT)
   mutate(Depth_bins = as.character(ifelse(Site %in% c(98,96,92,90) & Depth_m > 0.1, "BOT", Depth_m)),
          Depth_bins = as.character(ifelse(Site %in% c(88,50) & Depth_m > 9, "BOT", Depth_bins)),
          Depth_bins = as.character(ifelse(!Depth_bins %in% c("0.1", "BOT"), "Int", Depth_bins))
  ) |> 
  #zscores
  mutate(Temp_zs = scale(Temp_C),
         SC_zs = scale(SpCond_uScm),
         dexcess_zs = scale(D_excess),
         BIX_zs = scale(BIX),
         HIX_zs = scale(HIX),
         SUVA254_zs = scale(SUVA254),
         AT_zs = scale(A_T),
         DOC_zs = scale(DOC_mgL),
         NO3_zs = scale(NO3NO2_ugL),
         Chla_zs = scale(Chla_ugL)
         )

  
```



## singular breakpoint

example code for running a singular breakpoint

```{r}
### BIX as an example

# #fit normal linear regression 
# my.lm <- lm(BIX ~ Distance, data = chem)
# summary(my.lm)
# davies <- davies.test(my.lm, k=10)
# davies
# 
# #Run breakpoint
# my.seg <- segmented(my.lm,  seg.Z = ~ Distance,  psi = 2 )
# summary(my.seg)
# my.seg$psi
# slope(my.seg)
# 
# #fit breakpoint segments to distance data
# my.fitted <- fitted(my.seg)
# my.model <- data.frame(Distance = chem$Distance, BIX_fit = my.fitted)
# 
# #get breakpoint lines
# #NOTE that this line comes up chopy because its fitting to x axis values that aren't continous (have gaps between c and P sites; fixed later below)
# my.lines <- my.seg$psi[ , 2]
# 
# #plot
# chem |> 
#   ggplot(aes(x = Distance, y = BIX, fill = Depth_m))+
#   geom_point(shape = 21, size = 4) + #shape = 21,
#   # geom_smooth()+
#   theme_bw()+ theme(legend.position = "top", text = element_text(size = 18),
#                     panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
#   labs(x= "Distance from Stream (m)", y = "BIX", fill = "Depth (m)",
#        shape = "Site Type")+
#   scale_x_continuous(breaks = c(0, 500, 1000, 1500))+ 
#   scale_fill_gradient2(low = "red",  high ="blue",
#                        midpoint = 6,  guide = "colourbar", breaks = c(3,6,9, 15))+
#   #geom_smooth()+
#   geom_vline(xintercept = my.lines, linetype = "dashed", size = 1.5) +
#   geom_line(data = my.model, aes(x = Distance, y = BIX_fit), size = 1, inherit.aes = FALSE ) 



## get the quick stats by running on variables breakpoint
# #BIX
# my.lm <- lm(BIX ~ Distance, data = chem)
# davies <- davies.test(my.lm, k=10)
# davies
# my.seg <- segmented(my.lm,  seg.Z = ~ Distance,  psi = 2 )
# summary(my.seg)
# my.seg$psi
# slope(my.seg)


```



## Breakpoint function for MS figures

```{r}

run_breakpoint_analysis <- function(df, 
                                    response_var = NULL,
                                    response_var_label = NULL,
                                    fill_var = NULL,
                                    fill_var_label = NULL,
                                    title = NULL,
                                    depth_shape = FALSE  # TRUE/FALSE to map Depth_bins to shape
                                    ) {
  
  # Fit linear model
  my.lm <- lm(as.formula(paste(response_var, "~ Distance")), data = df)
  
  # Davies test
  davies <- davies.test(my.lm, k = 10)
  
  # Segmented model
  my.seg <- segmented(my.lm, seg.Z = ~ Distance, psi = 2)
  
  # Make regression lines
  dist_gradient <- data.frame(Distance = seq(0, max(df$Distance, na.rm = TRUE), by = 1))
  dist_gradient$fit <- predict(my.seg, newdata = dist_gradient)
  
  # Breakpoint + SE
  my.lines <- my.seg$psi[, 2]
  my.lines.SE <- my.seg$psi[, 3]
  
  # Build base plot
  p <- df |>
    ggplot(aes(x = Distance, y = .data[[response_var]], fill = .data[[fill_var]])) +
    geom_rect(
      aes(
        xmin = (my.lines - my.lines.SE),
        xmax = (my.lines + my.lines.SE),
        ymin = -Inf, ymax = Inf
      ),
      alpha = 0.1, fill = "gray", color = NA
    )
  
  # Conditionally map shape to Depth_bins or use fixed shape
  if (depth_shape) {
    p <- p +
      geom_point(aes(shape = Depth_bins), size = 4) +
            scale_shape_manual(
        values = c("0.1" = 21, "Int" = 22, "BOT" = 24),
        limits = c("0.1", "Int", "BOT"),
        name = "Depth"
      )
  } else {
    p <- p +
      geom_point(shape = 21, size = 4)
      #geom_jitter(shape = 21, size = 4, width = 20)
    
  }
  
  p <- p +
    labs(
      x = "Distance from Stream (m)",
      title = title,
      y = response_var_label,
      fill = fill_var_label
    ) +
    scale_x_continuous(breaks = c(0, 500, 1000, 1500)) +
     geom_vline(xintercept = my.lines, linetype = "dashed", size = 1.5) +
    geom_line(
      data = dist_gradient,
      aes(x = Distance, y = fit),
      size = 1,
      inherit.aes = FALSE
    ) +
    annotate("text",
             x = c(0, 189, 336, 1013),
             y = Inf,
             label = c("S1", "B1", "C1", "P1"),
             vjust = -0.5,
             size = 5) +
    coord_cartesian(clip = "off") +
    theme_bw() +
    theme(
      legend.position = "top",
      text = element_text(size = 14),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.background = element_blank(),
      legend.box.background = element_blank()
    )
  #change fill color ramp for filling by Depth vs other variables
    if (fill_var == "Depth_m") {
    p <- p + scale_fill_gradient2(
      low = "red", high = "blue",
      midpoint = 6, guide = "colourbar",
      breaks = c(3, 6, 9, 15)
    )
  } else {
    p <- p + scale_fill_distiller(palette = "Spectral", direction = -1)
  }
 
  list(
    plot = p,
    segmented_summary = summary(my.seg),
    segmented_slopes = slope(my.seg),
    davies_test = davies
  )
} # end of function


```


test that function is working, can toggle on and off inputs below
```{r}
run_breakpoint_analysis(chem_bp, response_var = "BIX", response_var_label = "BIX", title = "BIX function test",
                                        fill_var = "SpCond_uScm", fill_var_label = "Sp Cond", depth_shape = F)


```



#### Breakpoints for stream-res boundary ---------------------------------------------------------------------------------------------------------

Temp, SpCond, d-excess

```{r}
## Temp
temp_break <- run_breakpoint_analysis(chem_bp, response_var = "Temp_C", response_var_label = "Temp (C)",
                                      fill_var = "Depth_m", fill_var_label = "Depth (m)")

temp_break$plot
temp_break$davies_test

## Cond
SC_break <- run_breakpoint_analysis(chem_bp, response_var = "SpCond_uScm", response_var_label = "SC",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

SC_break$plot
SC_break$davies_test

## DO
DO_break <- run_breakpoint_analysis(chem_bp, response_var = "DOsat_percent", response_var_label = "DOsat",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

DO_break$plot
DO_break$davies_test


## d-excess
Dexcess_break <- run_breakpoint_analysis(chem_bp, response_var = "D_excess", response_var_label = "d-excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)")

Dexcess_break$plot
Dexcess_break$davies_test


## stack Sp cond and d-excess
Dexcess_break$plot / SC_break$plot



```


Z score SC and dexcess for main MS

```{r}

## Cond
SC_zs_break <- run_breakpoint_analysis(chem_bp, response_var = "SC_zs", response_var_label = "SC (z-score)", fill_var = "Depth_m", fill_var_label = "Depth (m)")

SC_zs_break$plot
SC_zs_break$davies_test
SC_zs_break$segmented_summary
SC_zs_break$segmented_slopes



## d-excess
Dexcess_zs_break <- run_breakpoint_analysis(chem_bp, response_var = "dexcess_zs", response_var_label = "d-excess (z-score)",fill_var = "Depth_m", fill_var_label = "Depth (m)")

Dexcess_zs_break$plot
Dexcess_zs_break$davies_test
Dexcess_zs_break$segmented_summary
Dexcess_zs_break$segmented_slopes



```





### Show changes in Temp, Sc, DO, and d-excess by date

```{r}

levels <- c("Temp_C", "SpCond_uScm", "D_excess")

levels_db <- c("22 May", "19 Jun", "11 Jul", "14 Aug", "16 Sep", 
               "30 Sep", "28 Oct", "17 Dec", "26 Feb", "16 Apr")

#By dates
chem_bp |> 
  dplyr::select(-Depth_bins) |> 
  pivot_longer(-c(1:7)) |>  
  # filter(!Site %in% c(50,88,90)) |> #if wanting to zoom in on S1 - C1 transition
  filter(name %in% levels) |> 
  mutate(Date1 = format(Date, "%d %b")) |> 
  ggplot(aes(x = Distance, y = value, fill = Depth_m))+
  geom_rect(aes(xmin = Dry_start, xmax = Dry_end, ymin = -Inf, ymax = Inf), 
            alpha = 0.1, fill = "gray", color = NA )+ #color gets rid of border
  geom_point(shape = 21, size = 4) + #shape = 21,
  #geom_smooth()+ # add ', warning=FALSE' to chunk to see plot
  facet_grid(factor(name, levels = levels)~factor(Date1, levels = levels_db), scales = "free_y")+
  theme_bw()+ theme(legend.position = "top", text = element_text(size = 14),
                    axis.text.x = element_text(angle = 45, hjust = 1), #can make angle = 45 and hjust = 1 to shift x label
                    panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(x= "Distance from Stream (m)", fill = "Depth (m)")+
  scale_x_continuous(breaks = c(0, 600, 1200))+ 
  scale_fill_gradient2(low = "red",  high ="blue",
                       midpoint = 6,  guide = "colourbar", breaks = c(3,6,9, 15))

# ggsave("../Figures/Temp_SC_Dexcess_by_Date_SI.png", height = 8, width = 11, units = "in")



```



### Check if this changes by days with dry streams or days with plunge points

Set up filtering for day groups

```{r}
### dry stream vs connected days 
dry_stream_days <- c(ymd("2024-06-19"), ymd("2024-07-11"), ymd("2024-08-14"), 
                ymd("2024-09-16"), ymd("2024-10-28") )

wet_stream_days <- c(ymd("2024-05-22"), ymd("2024-09-30"), ymd("2024-12-17"), 
                ymd("2025-02-26"), ymd("2025-04-16"))

### plunge v no plunge groups 
plungedays <- c(ymd("2024-05-22"), ymd("2024-06-19"), ymd("2025-02-26"), ymd("2025-04-16"))

nonplungedays <- c(ymd("2024-07-11"), ymd("2024-08-14"), ymd("2024-09-16"),
                    ymd("2024-09-30"),  ymd("2024-10-28"), ymd("2024-12-17"))


```

Plunge vs No plunge BP's

```{r}
## SC
plungedays_df <- chem_bp |> 
  filter(Date %in% plungedays)

SC_break_plunge <- run_breakpoint_analysis(plungedays_df, response_var = "SpCond_uScm", response_var_label = "SC",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)", title = "Plunge Days")

SC_break_plunge$davies_test


noplungedays_df <- chem_bp |> 
  filter(Date %in% nonplungedays)

SC_break_plungeno <- run_breakpoint_analysis(noplungedays_df, response_var = "SpCond_uScm", response_var_label = "SC",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)", title = "Non Plunge Days")

SC_break_plungeno$davies_test

SC_break_plunge$plot / SC_break_plungeno$plot


##D-excess
Dexcess_break_plunge <- run_breakpoint_analysis(plungedays_df, response_var = "D_excess", response_var_label = "d-excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)", title = "Plunge Days")

Dexcess_break_plunge$davies_test


Dexcess_break_plungeno <- run_breakpoint_analysis(noplungedays_df, response_var = "D_excess", response_var_label = "d-excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)", title = "Non Plunge Days")

Dexcess_break_plungeno$davies_test

Dexcess_break_plunge$plot / Dexcess_break_plungeno$plot


```

Dry vs Wet BP's

```{r}
## SC
drystream_df <- chem_bp |> 
  filter(Date %in% dry_stream_days)

SC_break_dry <- run_breakpoint_analysis(drystream_df, response_var = "SpCond_uScm", response_var_label = "SC",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)", title = "Dry Stream Days")

SC_break_dry$davies_test


wetstream_df <- chem_bp |> 
  filter(Date %in% wet_stream_days)

SC_break_wet <- run_breakpoint_analysis(wetstream_df, response_var = "SpCond_uScm", response_var_label = "SC",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)", title = "Connected Stream Days")

SC_break_wet$davies_test

SC_break_dry$plot / SC_break_wet$plot


##D-excess
D_break_dry <- run_breakpoint_analysis(drystream_df, response_var = "D_excess", response_var_label = "d-excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)", title = "Dry Stream Days")

D_break_wet <- run_breakpoint_analysis(wetstream_df, response_var = "D_excess", response_var_label = "d-excess",
                                    fill_var = "Depth_m", fill_var_label = "Depth (m)", title = "Connected Stream Days")

D_break_dry$plot / D_break_wet$plot



```


SI figure for plunge and dry stream days 

```{r}

##Plunge v no plunge
SC_break_plunge$plot / SC_break_plungeno$plot

Dexcess_break_plunge$plot / Dexcess_break_plungeno$plot

cowplot::plot_grid(SC_break_plunge$plot, Dexcess_break_plunge$plot , 
                                    SC_break_plungeno$plot, Dexcess_break_plungeno$plot,
                                     labels = c("a", "b", "c", "d"), ncol = 2)

#dry v wet 
SC_break_dry$plot / SC_break_wet$plot
D_break_dry$plot / D_break_wet$plot

cowplot::plot_grid(SC_break_dry$plot, D_break_dry$plot , 
                                    SC_break_wet$plot, D_break_wet$plot,
                                     labels = c("a", "b", "c", "d"), ncol = 2)


## all together now 
SI_plunge_dry_SC_dexcess_fig <-
  cowplot::plot_grid(SC_break_plunge$plot, Dexcess_break_plunge$plot , SC_break_dry$plot, D_break_dry$plot , 
                   SC_break_plungeno$plot, Dexcess_break_plungeno$plot, SC_break_wet$plot, D_break_wet$plot,
                                    # labels = c("a", "b", "c", "d", "e", "f", "g", "h"), 
                   labels = c("a", "c", "e", "g", "b", "d", "f", "h"), ncol = 4)


#ggsave("../Figures/Plunge_dry_SC_Dexcess_SI.png", SI_plunge_dry_SC_dexcess_fig, height = 8, width = 13, units = "in")


```





#### DOM Breakpoints --------------------------------------------------------------------------------------------------------------------

## Multiple DOM variables
across all samples

```{r}

#### DOC
doc_zt_break <- run_breakpoint_analysis(chem_bp, response_var = "DOC_zs", response_var_label = "DOC (z-score)",
                                        fill_var = "Depth_m", fill_var_label = "Depth (m)")

doc_zt_break$plot
doc_zt_break$davies_test
doc_zt_break$segmented_summary
doc_zt_break$segmented_slopes

#### BIX

bix_zt_break <- run_breakpoint_analysis(chem_bp, response_var = "BIX_zs", response_var_label = "BIX (z-score)",
                                        fill_var = "Depth_m", fill_var_label = "Depth (m)")

bix_zt_break$plot
bix_zt_break$davies_test
bix_zt_break$segmented_summary
bix_zt_break$segmented_slopes

#### HIX
hix_zt_break <- run_breakpoint_analysis(chem_bp, response_var = "HIX_zs", response_var_label = "HIX \n (z-score)",  fill_var = "Depth_m", fill_var_label = "Depth (m)")
hix_zt_break$davies_test
hix_zt_break$segmented_summary
hix_zt_break$segmented_slopes


#### AT
AT_zt_break <- run_breakpoint_analysis(chem_bp, response_var = "AT_zs", response_var_label = "Peak A:T \n (z-score)",  fill_var = "Depth_m", fill_var_label = "Depth (m)")
AT_zt_break$davies_test
AT_zt_break$segmented_summary
AT_zt_break$segmented_slopes



#### SUVA
SUVA_zt_break <- run_breakpoint_analysis(chem_bp, response_var = "SUVA254_zs", response_var_label = "SUVA254 \n (z-score)", fill_var = "Depth_m", fill_var_label = "Depth (m)")
SUVA_zt_break$davies_test
SUVA_zt_break$segmented_summary
SUVA_zt_break$segmented_slopes


cowplot::plot_grid(AT_zt_break$plot, SUVA_zt_break$plot, hix_zt_break$plot,
                                     labels = c("a", "b", "c"), ncol = 1)



```


Make MS breakpoint figure

```{r}

BP_ms <- cowplot::plot_grid(SC_zs_break$plot, Dexcess_zs_break$plot ,
                   doc_zt_break$plot, bix_zt_break$plot,
                                     labels = c("a", "b", "c", "d"), ncol = 1)


# ggsave("../Figures/BP_MS.png", BP_ms, height = 9, width = 7, units = "in")

```



## BIX to lots of responses for fill

```{r}

#stack
library(patchwork)
SC_break$plot / Dexcess_break$plot / bix_break$plot / doc_break$plot

#bix other fills
bix_zt_break_sc <- run_breakpoint_analysis(chem_bp, response_var = "BIX_zs", response_var_label = "BIX (z-score)",
                                      fill_var = "SpCond_uScm", fill_var_label = "SpCond")

bix_zt_break_dexcess <- run_breakpoint_analysis(chem_bp, response_var = "BIX_zs", response_var_label = "BIX (z-score)",
                                      fill_var = "D_excess", fill_var_label = "d-excess")

bix_zt_break_stage <- run_breakpoint_analysis(chem_bp, response_var = "BIX_zs", response_var_label = "BIX (z-score)",
                                      fill_var = "HPB_stage_cm", fill_var_label = "stream stage")

bix_bp_colorrampling <- (bix_zt_break$plot / bix_zt_break_sc$plot) | (bix_zt_break_dexcess$plot / bix_zt_break_stage$plot)

# ggsave("../Figures/BIX_BPs_colorramps.png", bix_bp_colorrampling, height = 8, width = 9, units = "in")



```




## Surf vs bottom breaks for SI fig

set up surface and max depth DFs

```{r}
##make surface only data frame
chem_surf <- chem_bp |>
  filter(Depth_m == 0.1)

##make max depth only data frame
chem_maxdepth <- chem_bp |>
  group_by(Site, Date) |>
  mutate(max_depth = if_else(Depth_m == max(Depth_m, na.rm = TRUE), "Max", "Not")) |>
  ungroup() |>
  filter(max_depth == "Max")

```


BIX breakpoints: all depths, surface, bottom To interpret stats outputs its easier to run chunk line by line

```{r}
#### BIX 

## BIX surface
bix_zt_surf_break <- run_breakpoint_analysis(chem_surf, response_var = "BIX_zs", response_var_label = "BIX (z-score)",
                                        fill_var = "Depth_m", fill_var_label = "Depth (m)")
bix_zt_surf_break$davies_test
bix_zt_surf_break$segmented_summary
bix_zt_surf_break$segmented_slopes

## BIX BOT
bix_zt_bot_break <- run_breakpoint_analysis(chem_maxdepth, response_var = "BIX_zs", response_var_label = "BIX (z-score)",
                                        fill_var = "Depth_m", fill_var_label = "Depth (m)")
bix_zt_bot_break$davies_test
bix_zt_bot_break$segmented_summary
bix_zt_bot_break$segmented_slopes


#plot breakpoints
bix_zt_break$plot / bix_zt_surf_break$plot / bix_zt_bot_break$plot


```

DOC breakpoints: all depths, surface, bottom

```{r}
#### DOC 

## DOC surface
doc_zt_surf_break <- run_breakpoint_analysis(chem_surf, response_var = "DOC_zs", response_var_label = "DOC (z-score)",
                                        fill_var = "Depth_m", fill_var_label = "Depth (m)")
doc_zt_surf_break$davies_test
doc_zt_surf_break$segmented_summary
doc_zt_surf_break$segmented_slopes

## DOC BOT
doc_zt_bot_break <- run_breakpoint_analysis(chem_maxdepth, response_var = "DOC_zs", response_var_label = "DOC (z-score)",
                                        fill_var = "Depth_m", fill_var_label = "Depth (m)")
doc_zt_bot_break$davies_test
doc_zt_bot_break$segmented_summary
doc_zt_bot_break$segmented_slopes


```

DOC bix for diff depths - SI fig

```{r}
##plot them all
DOM_surf_maxdepth_BPs <- cowplot::plot_grid(doc_zt_surf_break$plot,  bix_zt_surf_break$plot,
                                            doc_zt_break$plot, bix_zt_break$plot,  
                                           doc_zt_bot_break$plot, bix_zt_bot_break$plot,
                                     labels = c("a) DOC surface", "b) BIX surface",
                                                "c) DOC all", "d) BIX all",
                                                "e) DOC max", "f) BIX max"), ncol = 2)

DOM_surf_maxdepth_BPs

# ggsave("../Figures/DOM_surf_bot_BPs.png", DOM_surf_maxdepth_BPs, height = 8, width = 9, units = "in")


```




## NO3 and Chla breaks

NO3 all depths

```{r}
####NO3
no3_zt_break <- run_breakpoint_analysis(chem_bp, response_var = "NO3_zs", response_var_label = "NO3 \n (z-score)", fill_var = "Depth_m", fill_var_label = "Depth (m)")

no3_zt_break$plot
no3_zt_break$davies_test
no3_zt_break$segmented_summary
no3_zt_break$segmented_slopes


```


Make SI figure for NO3 and other DOM breakpoints

```{r}
DOM_NO3_SI_break_plots <- 
  cowplot::plot_grid( AT_zt_break$plot, SUVA_zt_break$plot,
                      hix_zt_break$plot,  no3_zt_break$plot, 
                                     labels = c("a", "b", "c", "d"), ncol = 1)

DOM_NO3_SI_break_plots

# ggsave("../Figures/DOM_NO3_SI_BP.png", DOM_NO3_SI_break_plots, height = 9, width = 9, units = "in")



```





NO3 and chla surface

```{r}
#### NO3 surface
no3_zt_break_surf <- run_breakpoint_analysis(chem_surf, response_var = "NO3_zs", response_var_label = "NO3 (z-score)",
                                        fill_var = "Depth_m", fill_var_label = "Depth (m)")

no3_zt_break_surf$plot
no3_zt_break_surf$davies_test
no3_zt_break_surf$segmented_summary
no3_zt_break_surf$segmented_slopes


#### chla surface
chla_zt_break_surf <- run_breakpoint_analysis(chem_surf, response_var = "Chla_zs", response_var_label = "Chl-a (z-score)",
                                        fill_var = "Depth_m", fill_var_label = "Depth (m)")

chla_zt_break_surf$plot
chla_zt_break_surf$davies_test
chla_zt_break_surf$segmented_summary
chla_zt_break_surf$segmented_slopes



##make figure 
chla_break_plots <- cowplot::plot_grid(no3_zt_break_surf$plot, chla_zt_break_surf$plot, 
                                       doc_zt_surf_break$plot,  bix_zt_surf_break$plot, 
                                     labels = c("a", "b", "c", "d"), ncol = 1)

chla_break_plots

#ggsave("../Figures/Chla_breakpoints_surf_SI.png", chla_break_plots, height = 9, width = 7, units = "in")





```



## regressions to sensors

CQ plot

```{r}
#get flowmate data
sampling_dates <- c(ymd("2024-05-22"), ymd("2024-06-19"), ymd("2024-07-11"), ymd("2024-08-14"), ymd("2024-09-16"),
                    ymd("2024-09-30"), ymd("2024-10-28"), ymd("2024-12-17"), ymd("2025-02-26"), ymd("2025-04-16"))

flowmate_edi <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/454/9/0e7fe16623a1ad2a67774c23ce8a29d8")

hpb_flowmate <- flowmate_edi |> 
  mutate(Date = as.Date(DateTime)) |> 
  filter(Date %in% sampling_dates) |> 
  dplyr::select(Date, Flow_cms)

```


```{r}
chem_bp |> left_join(hpb_flowmate, by = "Date") |> 
  filter(Site_code == "S1") |> 
  dplyr::select(Date, Flow_cms, DOC_mgL, BIX, HIX, NO3NO2_ugL ) |> 
  pivot_longer(-c(1:2)) |> 
  ggplot(aes(x = Flow_cms, y = value, color = Date))+
  geom_point()+
  facet_wrap(~name, scales = "free_y")+
  theme_bw()


```




```{r}
# chem_bp |> 
#   mutate(Site_Type = ifelse(Site_code %in% c("S1", "S2"), "Stream", NA),
#          Site_Type = ifelse(Site_code %in% c("B1", "B2"), "Backwater", Site_Type),
#          Site_Type = ifelse(Site_code %in% c("C1", "C2"), "Cove", Site_Type),
#          Site_Type = ifelse(Site_code %in% c("P1", "P2"), "Pelagic", Site_Type)
#          ) |> 
#   ggplot(aes(x = SpCond_uScm, y = BIX, shape = Site_Type, color = Date))+
#   geom_point()



```
